Psyc 601 Multivariate Statistics Project

### 2013-08-28 writeups/proposal/proposal.tex
foo
Former-commit-id: bbd9b39487b08d18524a02dfdfee377259a202fe [formerly 80f2178267c5ba456ae4cc6616ed4022a5aa58e4]
Former-commit-id: f5c86f73a91d0af415f1ec2c8e7bcd79541f55f3

FIXME: \cite{Godin2013}
This paragraph will discuss the topics-based approach.

FIXME: Paragraph that no vector-based models have been attempted for Twitter.
Bayesian model can be improved by customizing predicitons based on past hashtag use.

FIXME: ditch this paragraph?
To our knowledge, no model of user-created hashtags when composing tweets on Twitter has yet been built.
There are two likely reasons for this:
[1] Modeling user-chosen hashtags and building hashtag-recommendation systems has only explored over the past few years, and
[2] Twitter is a large database of human-generated content, spanning a wide range of topics, containing a sizeable set of user-generated hashtags, so hashtag prediction for Twitter is a relatively hard problem.
We are nonetheless curious if declarative memory retrieval models can be applied to this domain to generate useful hashtag predictions.
The domain shares many similarities with environments where hashtag-recommendation systems have already been deployed:
Google+'s microblogging service is essentially Twitter, but focused on a specific subset of users and topics.
The StackExchange sites are also built from a large database of human generated content, where users have chosen specific tags to associate with posts.
So we will evaluate how well an ACT-R-inspired Bayesian declarative memory model and a new vector-based model predict a user's chosen hashtag when composing a tweet on Twitter.

>>>>

A topics-based approach was used by \textcite{Godin2013}, where the model predicted the most likely set of topics that were associated with a tweet, and tested if these topics matched the tweet by using human raters.
The latent topics for the colleciton of tweets were generated using a Latent Dirichlet Allocation statistical methodology.
This is similar to singular value decomposition in that the most likely hidden dimensions (i.e., topics) of the data are discovered by searching for reduced representations of the data that cover most of the variance. 
Human graders rated how well the models generated topics for a tweet related to the content in the tweet.
The LDA model performed well, and generated relevant topics.

However, a topic-generating model is not performing the same task as a hashtag-prediction model.
The authors argued that these general topics could be used to categorize the post, in addition to the hashtags already used by the author in the post.
It is unclear though if a user would want to label their tweet with these general topics.
Perhaps they would rather use more specific hashtags to be more precise in how they label the tweet and how that tweet is associated with tweets within the rest of the community.
In fact, they have already chosen to be more precise in the specific labels that they use for the tweet, since they chose to label the tweet with specific hashtags instead of general topics.
Now it is certainly the case that the topics generated by a latent-based model may be similar, if not identical, to the hashtags used in the post.
However, it seems a bit indirect to build a model that predicts general topics, since embedded in the data are the exact ``topics'' (i.e., hashtags) that each user chose to use for each tweet.
Why not take those topics already present in the data and build a model that produces them?
In other words, the topics data are already in the tweet, so why not use them?

\subsubsection{State of Research on StackOverflow and Twitter}

The current state of research on Twitter hashtag prediction is that a mix of co-occurrence and latent-based models have been used to predict hashtags, and the results are encouraging.
However, there has been very little research on how a specific user's past hashtag use should influence the model's prediction when that user is composing a tweet.
\textcite{Kywe2012} did take a user-centered view with their model, but the hashtags predicted by the user and content were analyzed separately and then the top from each group were combined for prediction.
The ACT-R declarative memory retrieval model takes a combined approach, and shows how past user behavior and current context combine to produce the most likely retrieval.
Each component is summed together to generate a total activation, and then the likelihodd of chunks are ranked by that summed activation.
This technique allows for a more natural combination of the two components, where weights can be assigned to each component depending on how strongly each term predicts performance.
This type of weighted additive model that combines past user behavior and current tweet context to predict the most likely hashtags has not yet been explored for Twitter.

Also, a vector-based retrieval model has not yet been tested on Twitter.
One of the advantages of vector-based models is that word order can be easily incorporated into the single co-occurrence representation \parencite{Jones2007}.
It seems quite likely that the few words used just before a hashtag embedded in a tweet will be highly predictive of the hashtag chosen.
To our knowledge, only bag-of-words models have been tested on Twitter to predict hashtags, where co-occurrance between words and hashtags are computed independently from where the word is located in the tweet.
It will be interesting to see how much (if at all) accuracy improves when incorporating word-order information into the model.
